code_revisit_2

### 2021-01-27 code_revisit2.Rmd
Add the summary of scenario 1 and 2

## Scenario 1 
- For the obs with more than one jobs, and one of the job has na in hours_work, we can not calculate the weighted average.
- For the obs with only one job and the hours_work is na, we only will use the rates directly.
- Calculate the weighted hours normally

```{r}
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = n_distinct(rate_per_hour))

eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))
```

@@

```{r}
wages_weighted %>% filter(id == 506)
```
```{r}
wages_tidy_sc1 <- wages_weighted_sc1 %>%
  filter(!is.na(wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 300,
                        NA, wages))
```

@@

ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = wages, 
                group = id), 
            alpha = 0.1)

@@

```{r}
set.seed(2020)
ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = wages,
                group = id)) +
  facet_sample()
```
```{r features}
wages_feature_sc1 <- wages_sc1 %>%
  features(wages, 
           feat_three_num
           )
```

@@

p1+p2+p3

@@

## SCENARIO 2
So, let us try scenario 2, by exclude the observation with more than one job with NA in hours_work.

@@

```
```{r}
wages_tidy_sc2 <- wages_weighted_sc2 %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 300,
                        NA, wages))
```

```{r}
wages <- as_tsibble(x = wages_tidy_sc2,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
ggplot(wages) +
  geom_line(aes(x = year,
                y = wages,
                group = id), 
            alpha = 0.1)
```

```{r}
set.seed(2021)
ggplot(wages, aes(x = year,
                y = wages,
                group = id)) +
  geom_line() +
  facet_sample()
```

```{r features}
wages_feature<- wages %>%
  features(wages, 
           feat_three_num
           )
```

```{r p1}
p4 <- ggplot(wages_feature) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

@@

wages_feature_long <- wages_feature %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

p5 <- ggplot(wages_feature_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r p3}
p6 <- ggplot(transform(wages_feature_long,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

@@

p4+p5+p6

@@

>>>>

## SCENARIO 1 
- For the obs with more than one jobs, and one of the job has na in hours_work, we can not calculate the weighted average.
- For the obs with only one job and the hours_work is na, we only will use the rates directly.
- Calculate the weighted hours normally

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = n_distinct(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))
```

@@

```{r}
wages_weighted_sc1 %>% filter(id == 506)
```
```{r}
wages_tidy_sc1 <- wages_weighted_sc1 %>%
  filter(!is.na(wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 30,
                        NA, wages))
```

@@

sc1_plot <- ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = wages, 
                group = id), 
            alpha = 0.1) +
  ggtitle("Scenario 1")

sc1_plot

@@

```{r}
set.seed(2020)
ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = wages,
                group = id)) +
  facet_sample()

```
```{r features}
wages_feature_sc1 <- wages_sc1 %>%
  features(wages, 
           feat_three_num
           )
```

@@

sc1 <- p1+p2+p3
sc1

@@

## SCENARIO 2

So, let us try scenario 2, by exclude the observation with more than one job with NA in hours_work.

@@

```
```{r}

wages_weighted_sc2 %>% filter(id == 506)
```

```{r}
wages_tidy_sc2 <- wages_weighted_sc2 %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 30,
                        NA, wages))
```

```{r}
wages_sc2 <- as_tsibble(x = wages_tidy_sc2,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
sc2_plot <- ggplot(wages_sc2) +
  geom_line(aes(x = year,
                y = wages,
                group = id), 
            alpha = 0.1) +
  ggtitle("Scenario 2")
```

```{r}
set.seed(2021)
ggplot(wages_sc2, aes(x = year,
                y = wages,
                group = id)) +
  geom_line() +
  facet_sample()
```

```{r features}
wages_feature_sc2<- wages_sc2 %>%
  features(wages, 
           feat_three_num
           )
```

```{r p1}
p4 <- ggplot(wages_feature_sc2) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

@@

wages_feature_long_sc2 <- wages_feature_sc2 %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

p5 <- ggplot(wages_feature_long_sc2) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r p3}
p6 <- ggplot(transform(wages_feature_long_sc2,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

@@

sc2 <- p4+p5+p6
sc2

@@

## SCENARIO 3 (modified SC1)

In this scenario, we treated the observation with the wages higher than $100 as NA

```{r}
wages_tidy_sc3 <- wages_weighted_sc1 %>%
  filter(!is.na(wages)) %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 100,
                        NA, wages))
```

```{r}
wages_sc3 <- as_tsibble(x = wages_tidy_sc3,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
ggplot(wages_sc3) +
  geom_line(aes(x = year,
                y = wages,
                group = id), 
            alpha = 0.1)
```

```{r}
set.seed(2021)
ggplot(wages_sc3, aes(x = year,
                y = wages,
                group = id)) +
  geom_line() +
  facet_sample()
```

## SCENARIO 4 (modified SC2)

```{r}
wages_tidy_sc4 <- wages_weighted_sc2 %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  mutate(wages = ifelse(wages > 100,
                        NA, wages))
```

```{r}
wages_sc4 <- as_tsibble(x = wages_tidy_sc4,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
ggplot(wages_sc4) +
  geom_line(aes(x = year,
                y = wages,
                group = id), 
            alpha = 0.1)
```

```{r}
set.seed(2021)
ggplot(wages_sc4, aes(x = year,
                y = wages,
                group = id)) +
  geom_line() +
  facet_sample()
```

## SUMMARY

### The original tidy wages

```{r}
ori <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv") %>%
  mutate(mean_hourly_wage = ifelse( mean_hourly_wage > 30,
                        NA, mean_hourly_wage))
```

```{r}
fig_all_transp <- ggplot(ori) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = CASEID_1979),
            alpha = 0.1) +
  ggtitle("Original data")
```

```{r wages_tsibble}
wages_ori <- as_tsibble(x = ori,
                    key = CASEID_1979,
                    index = year,
                    regular = FALSE)
```

```{r}
wages_obs <- wages_ori %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
```

### sc1 + sc2 + original tidy employment data

```{r}
sc1_plot + sc2_plot + fig_all_transp
```

Number of observations

```{r}
nrow(wages_ori) #regular mean
nrow(wages_sc1) #weighted mean omit id with NA in the working hours
nrow(wages_sc2) #weighted mean omit obs with NA in the working hours
```

Number of ids

```{r}
nrow(wages_obs)
nrow(wages_feature_sc1)
nrow(wages_feature_sc2)
```
